### 2017-06-24 submissions/ohbm_2017/Poster/poster.tex
Big picture plots

\pdfminorversion=4
\documentclass[portrait,a0,final]{a0poster} % a0poster class sets the paper size
\usepackage[RGB,A0]{ScientificPosterTemplate} % RGB, Coated, Uncoated
\usepackage[english]{babel}
\usepackage[pdftex]{graphicx}
\usepackage{epstopdf} % so that ''pdflatex'' accepts .eps files
\usepackage{titlesec} % For changing the font on chapters, sections, etc.
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{tabularx}
\usepackage{lipsum} % Lorem ipsum generator
\usepackage{floatrow}

\definecolor{bleu}{RGB}{0,140,189}%

%% Section font size reduction for a0 posters %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \titleformat{\section}{\large\bfseries\sffamily\color{posterBlack}}{\textcolor{posterBlack}{\thesection}}{1em}{} % Text size for a1 posters
% \titleformat{\section}{\large\bfseries\sffamily}{\thesection.}{1em}{} % Text size for a1 posters with a dot after the incremental number
\titleformat{\section}{\huge\bfseries\sffamily\color{bleu}}{\textcolor{bleu}{\thesection.}}{1em}{} % Text size for a1 posters with a dot after the incremental number

%% Safe alternative math font (optional)
% \usepackage{fouriernc}

%% Teach hyphenation for alien words
% \hyphenation{op-tical net-works semi-conduc-tor}

%% Font sizes for a0poster are
%\tiny
%\scriptsize
%\footnotesize
%\small
%\normalsize
%\large
%\Large
%\LARGE
%\huge
%\Huge
%\veryHuge
%\VeryHuge
%\VERYHuge

%% Colors from ScientificPosterTemplate-package
% posterBlack
% posterGray
% posterYellow
% posterOrange
% posterRed
% posterFuchsia
% posterPurple
% posterBlue
% posterTurquoise
% posterGreen
% posterLightGreen
% You can use \textcolor{<posterColor>}{<your text>) to change the colors of text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\thispagestyle{empty} % Removes the page number

\begin{minipage}[t]{0.98\linewidth} % The first minipage for the logo & title
\vspace{0pt} % A trick to align the parallel minipages on top

\vspace{0.008\linewidth} % Increase the top margin

\begin{minipage}[t]{0.95\linewidth} % title
\vspace{0pt} % Alingns the parallel minipages on top

%% Title: use \baselinestretch to change linespacing, \textit{} for italic text, \textcolor for colored text
% More conservative title, upright and black
{\renewcommand{\baselinestretch}{0.8} % Changes the baseline skip smaller for the title
\VeryHuge{\bfseries{\textsf{\textcolor{bleu}{Predictive models based on
functional connectivity: a benchmark across multiple datasets}}}} % Text size for a1 posters
\par} % <- for \baselinestretch

\vspace{0.02\linewidth} % Empty space after the title

\Large{\textsf{\bfseries{Kamalaker DADI$^{1}$, Alexandre ABRAHAM$^1$,
Mehdi RAHIM$^1$, Darya CHYZHYK$^1$, Bertrand THIRION$^1$,
Ga{\"e}l VAROQUAUX$^1$}}} % Text size for a1 posters

\vspace{0.01\linewidth}
\Large{\textcolor{posterBlack}{\textsf{%\bfseries{
            $^1$Parietal team, Neurospin -- CEA, INRIA, Paris-Saclay, France%}
}}}
%
\end{minipage}
\begin{minipage}{0.76\paperwidth}
\vspace{-0.035\linewidth}
\begin{flushright}
\includegraphics[height=15mm, angle=0]{figures/logos/parietal_web_page.png}
\qquad \quad
\includegraphics[height=15mm, angle=0]{figures/logos/neurospin.jpg}
\qquad \quad
\includegraphics[height=26mm, angle=0]{figures/logos/logo-inria2.jpg}
\end{flushright}
\end{minipage} % No empty line before the second begin!
%\hspace{0.01\linewidth}
%\begin{minipage}{0.95\linewidth} % logo
%    \hfill
%\vspace{0pt} % Alingns the parallel minipages on top
%\begin{flushright}
%\includegraphics[width=0.20\linewidth,
%angle=0]{figures/logos/parietal_web_page.png}
%\end{flushright}

%\end{minipage} % no empty line before the next begin

\end{minipage}

%% Two columns

% Space according to the visual identity guidelines...
\vspace{0.03\linewidth}

% Centering helps in placement
\centering
\Large % Text size for a1 posters
\sffamily % font sans serif

\begin{minipage}{0.98\linewidth}

\begin{minipage}[t]{0.48\linewidth}
\setlength{\parindent}{10mm} % Paragraph indent

\section{Introduction}
\begin{itemize}
    \item Psychiatry and psychology are based on assessing individuals traits
        characterized through \textbf{behavioral testing} and
        \textbf{questionnaires}.
    \item \textbf{Imaging of brain activity} measures the physiological
        differences that underlie these \textbf{psychological
        variations}\cite{miller}.
    \item Introduction of \textbf{automated pipeline}\cite{abraham}.
        \begin{itemize}
            \item  Learning the \textbf{link} across individuals.
            \item Large cohorts of functional magnetic resonance images at
                rest \textbf{(Rest-fMRI)}.
        \end{itemize}
    \item We present the \textbf{outline} of this pipeline
        and how we used it to \textbf{draw best practices} from the
        application of pipeline on
        various problems.
\end{itemize}

\vspace{0.3\sectionspace}
\section{A connectome classification pipeline}
%We applied the pipeline on five datasets to i) determine the steps to obtain
%the best
%prediction, and ii) predict phenotypic information with good accuracy.

%\vspace{\figurespace}
\begin{center}
\includegraphics[width=1\linewidth]{figures/functional_connectivity_pipeline.pdf}
\end{center}
\captionof{figure}{\textbf{Our pipeline consists of three main steps}: defines regions
    from rest fMRI,
    builds connectomes from time series signals extracted from these
    regions of interests, and compares connectomes across subjects using
machine learning. \label{fig_1}}
%\vspace{\figurespace}

\vspace{0.0991\sectionspace}
\section{Rest fMRI datasets}

\begin{center}
    \begin{tabular}{ccc}
        \rowcolor{gray!50}
        Dataset & Subjects & Clinical question \\
        \hline\\[-4mm]
        COBRE & $65/77$ & Schizophrenia vs Control \\
        \rowcolor{gray!13}
        ADNI & $40/96$ & AD vs MCI  \\
        ADNIDOD & $89/78$& PTSD vs Control \\
        \rowcolor{gray!13}
        ACPI & $62/64$ & Marijuana use vs Control \\
        ABIDE & $402/464$ & Autism vs Control
    \end{tabular}
\end{center}
\captionof{table}{Description of the \textbf{five rs-fMRI
    datasets} used.
    AD - Alzheimer's Disease, MCI - Mild
    Cognitive Impairment,
    PTSD - Post Traumatic Stress Disorder.}
%\vspace{\tablespace}

\section{Pipelining choices}
    %\begin{enumerate}
    %   \item Defining Brain ROIs
    %   \item Connectivity parameterizing
    %   \item Supervised Learning: Classifiers
    %\end{enumerate}
%\vspace{\figurespace}
\begin{center}
\includegraphics[width=0.7\linewidth]{figures/only_pipeline_choices2.pdf}
\end{center}
\captionof{figure}{{\bfseries\sffamily 1 -- Defining Brain ROIs} can be done using
    three pre-defined atlases: two anatomical - Automated Anatomical Labeling
    (AAL), Harvard Oxford and one functional - Bootstrap Analysis
    of Stable Clusters (BASC). Four region learning methods: two Linear
    decomposition models - Group Independent Component Analysis (Group ICA),
    Online Dictionary Learning and two Clustering models - KMeans,
    Agglomerative with Ward criterion.
    {\bfseries\sffamily 2 -- Parameterizing functional connectivity
    between ROIs}: correlation, partial correlation or tangent space embedding
    {\bfseries\sffamily 3 -- Supervised learning}: Classifiers, a
classification
    model is built to predict groups with two linear classifiers, SVC
($\ell_{1}$
    or $\ell_{2}$ penalization) and Ridge. Definition of brain regions (below). \label{fig_2}}
%\vspace{\figurespace}

%\begin{centering}
\begin{minipage}{0.38\linewidth}
\begin{center}
\includegraphics[width=2.5\linewidth]{figures/rois2.pdf}
%\captionof{figure}{Definition of Brain regions. \label{fig_3}}
\end{center}
\end{minipage}
%\end{centering}
%\vspace{\figurespace}
%\begin{minipage}[t]{0.35\linewidth}
%        Here is the text
%          \end{minipage}\hfill
%            \begin{minipage}[b]{0.55\linewidth}
%                    \includegraphics[width=40cm]{figures/rois.pdf}
%                        \captionof{figure}{A wonderful picture}
%                          \end{minipage}
%

% The end of the first column and the start of the second
\end{minipage} % no empty line before the next ''\begin''
\hspace{0.02\linewidth} % Middle margin
\begin{minipage}[t]{0.48\linewidth}
\setlength{\parindent}{10mm} % Paragraph indent

\section{Results}
\subsection{Prediction scores in AUC}
\begin{tabularx}{\textwidth}{p{7.5cm}|p{5.5cm}|p{5.5cm}|p{5.5cm}|p{5.5cm}|p{5.5cm}}
    \rowcolor{gray!50}
    \centering{Accuracy} & \centering{COBRE} & \centering{ADNI} &
    \centering{ADNIDOD} & \centering{ACPI} & ABIDE \\
    \hline
    \rowcolor{gray!13}
    \centering{Median}& \centering{$87.1\%$} & \centering{$76.2\%$} &
    \centering{$67.1\%$} & \centering{$56.4\%$} & $70.2\%$\\
    % \rowcolor{gray!13}%
        {\centering $5^{th}$ Percentile} & \centering{$74.3\%$} &
    \centering{$62.3\%$} & \centering{$53.8\%$} & \centering{$41.4\%$} & $64.9\%$\\
    \rowcolor{gray!13}
    {\centering $95^{th}$ Percentile} & \centering{$94\%$} &
    \centering{$89.1\%$} & \centering{$77.2\%$} & \centering{$68.7\%$} & $73.8\%$\\
    \hline
\end{tabularx}
\captionof{table}{\textbf{Median, 5th percentile and 95th percentile of
        accuracy scores in AUC over cross-validation folds ($n=100$) across
        multi rs-fMRI datasets.} Functional brain atlases learned using Online
        Dictionary Learning\cite{arthur} with $60$ resting state networks
        and spliting networks to regions are used in the prediction pipeline
        with Tangent based connectivity matrix parameterization\cite{varoquaux} and
        Support Vector Classifier with $\ell{_2}$ penalization (SVC-$\ell{_2}$).}

\subsection{Choice of classifier, connectivity and brain atlas}
%\vspace{\figurespace}
\begin{center}
\begin{minipage}[t]{0.48\linewidth}
\begin{center}
\includegraphics[width=0.95\linewidth]{figures/impact_plot_classifier.pdf}\\\normalsize
(a) classifier
\end{center}
\end{minipage}
\hspace{0.02\linewidth} % Middle margin
\begin{minipage}[t]{0.48\linewidth}
\begin{center}
\includegraphics[width=0.87\linewidth]{figures/impact_plot_measure.pdf}\\\normalsize
(b) connectivity parameterization
\end{center}
\end{minipage}
\end{center}
%\captionof{figure}{The impact of classifier (a), connectivity
%parameterization (b)
%            choices in functional connectivity prediction pipeline over
%                    diverse tasks. \label{fig_3}}
%\vspace{\figurespace}

%\subsection{Choice of atlas}
%\vspace{\figurespace}
\begin{minipage}[t]{0.98\textwidth}
\begin{center}
\includegraphics[width=0.8\linewidth]{figures/impact_plot_atlas.pdf}\\
\normalsize (c) brain atlases with optimal choices in dimensionality
-- BASC (122 networks to regions) -- Ward (120) --
K-Means (120 parcellations) --
GroupICA (80 networks to regions) -- DictLearn (60 networks to regions)
%\vspace{\figurespace}
\end{center}
\end{minipage}
\captionof{figure}{The \textbf{impact of choices} in functional connectivity prediction
    pipeline over diverse tasks.
    (*) indicates the optimal option.
        \label{fig_3}}

%\vspace{\sectionspace}
\section{Optimal choices}
\begin{itemize}
    %\item Functional-Connectivity based pipeline across diverse clinical targets
    \item Systematic exploration of choices at each step
    \item Definition of Brain ROIs - Linear Decomposition models
    \item Connectivity Parameterization - Tangent Space Embedding
    \item Classification with Support Vector Classifier - $\ell{_2}$
        penalization.
\end{itemize}

\nocite{*}
%\vspace{\sectionspace}
\begin{spacing}{0.6}
\bibliographystyle{naturemag}
\bibliography{references}
\end{spacing}

\begin{spacing}{0.7}
\begin{minipage}[c]{0.675\linewidth}
    \begin{center}
\vspace{0.1\linewidth}
\textbf{\large Acknowledgments:}\\
\vspace{0.02\linewidth}
\large{ANR-11-BINF-0004\_NiConnect and CATI project}.
%\vspace{0.02\linewidth}
%\textcolor{posterGray}{\hrule}
%\vspace{0.02\textwidth}
%See Also Posters: \textbf{$\#1656$, $\#1829$, $\#3934$}
\end{center}

\end{minipage}
\end{spacing}

\begin{minipage}[c]{0.995\linewidth}
\vspace{0pt}
\begin{flushright}
\vspace{-0.03\linewidth}
\includegraphics[height=64mm, angle=0]{figures/logos/nilearn-logo.png}
\end{flushright}
\end{minipage}

\end{minipage}
\end{minipage} % minipage for the two columns (minipages)

\vfill % Fill the free space until the footer minipages

%\begin{minipage}{0.98\textwidth} % Minipages for the footers

%\large % Text size for footers

%\begin{minipage}[c]{0.475\linewidth}% Footer #1
%\vspace{0pt}
%\begin{center}
%\textcolor{posterGray}{\hrule}
%\vspace{0pt}
%\includegraphics[height=17mm, angle=0]{figures/logos/parietal_web_page.png}
%\qquad \quad
%\includegraphics[height=17mm, angle=0]{figures/logos/neurospin.jpg}
%\qquad \quad
%\includegraphics[height=29mm, angle=0]{figures/logos/logo-inria2.jpg}
%\end{center}
%\end{minipage} % No empty line before the second begin!
%\hspace{0.01\linewidth}
%\begin{minipage}[c]{0.575\linewidth} % Footer #2
%    \begin{center}
%\vspace{0pt}
%\textbf{\sffamily Acknowledgments:}\\
%ANR-11-BINF-0004\_NiConnect and
%CATI project.
%\end{center}
%\end{minipage}

%\end{minipage}

\vspace*{0.03\linewidth} % Increase the bottom margins

\end{document}